---
title: "Seattle Housing Market Analysis"
author: "Juliana Franco Correa"
date: "r Sys.Date()"
output: html_document
params:
   zipcode: 98052
---
# ================================================
# Practice Challenge 2: Reproducible Reports with R Markdown
# ================================================
# Instructions for a new file: report.Rmd

# R Markdown Setup
```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(DT)
library(knitr)
library(broom)
library(lubridate)
cleaned_house_data <- read_csv("cleaned_house_data.csv")
```

# Step 1: Generate two plots 
# (scatter plots for sqft_living vs price and bedrooms vs price)
```{r}
plota <- ggplot(cleaned_house_data, aes(x = sqft_living, y = price)) +
    geom_point(color = "darkred") +
    labs(title = "Price vs Square Footage", x = "Square Footage", y = "Price") + 
    theme_minimal()
plotb <- ggplot(cleaned_house_data, aes(x = bedrooms, y = price)) +
    geom_point(color = "darkred") +
    labs(title = "Price vs Bedrooms", x = "Bedrooms", y = "Price") + 
    theme_minimal()

plota
plotb
```

# Step 2: Create and format one table for summarized price by zipcode
```{r}
price_zipcode <- cleaned_house_data %>%
    group_by(zipcode) %>%
    summarise(avg_price = mean(price, na.rm = TRUE)) %>%
    arrange(desc(avg_price))
knitr::kable(price_zipcode)
```

# Step 3: Add narrative explanations (text chunks)
```{r}
# "The above scatter plots illustrate the relationship between price and two key features in housing: bedrooms and square footage."
# "The relationship between price and square footage appears to be more linear, eventhough it is not a perfect fit. However, the relationship between bedrooms and price is not quite clear."
# "On the other hand, the table summarizes the average price by zipcode."
```

# ================================================
# Practice Challenge 3: Interactive and Dynamic Reporting
# ================================================

# Step 1: Create an interactive plot with ggplotly()
```{r interactive_plot}
ggplotly(plota)
```
# Step 2: Add an interactive DT table with summary by zipcode
```{r interactive_table}
datatable(price_zipcode)
```

# Step 4: Filter house_data based on zipcode parameter
```{r filter_by_param}
filtered_data <- cleaned_house_data %>%
    filter(zipcode == params$zipcode)   
summary(filtered_data)
```

# Step 5: Knit report to HTML
# < PERFORM IN TERMINAL >