---
title: "Interactive Sales Report"
author: "Coursera Learner"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_caption: yes
params: 
    category: "Toys" # Can be configured to different parameter
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Background Code

The code below shows the associated libraries and loading of the datasets.
This section also produces a simple boxplot for units sold per region.
```{r}
# Load required libraries
library(dplyr)      # Data manipulation
library(ggplot2)    # Static plots
library(plotly)     # Interactive plotting
library(DT)         # Interactive tables
library(lubridate)  # Date handling
library(readr)      # Reading data

# Load and parse dataset
sales_data <- read_csv("sales_data.csv")
sales_data$Date <- as.Date(sales_data$date)

```

## Activity 1: Interactive Plots
### Steps:
- Use ggplot2 to create a basic plot
- Wrap it with plotly::ggplotly()

### Example
```{r}
# plot_qty <- ggplot(example_data, aes(x = location , y = sales)) +
#   geom_boxplot() +
#   labs(title = "Sales By location")  # Add context in title
# ggplotly(plot_qty)  # Make plot interactive

```

This section shows how to create interactive plots.

## Try it Out #1: Create a box plot of Units Sold over by Region and make it interactive.
### Steps:
- Use ggplot2 to plot Units_sold by Region
- Use geom_boxplot to visualize distribution
- Wrap the plot in ggplotly()

```{r}

plot_qty <- ggplot(sales_data, aes(x = region, y = units_sold)) +
  geom_boxplot() +
  labs(title = "Units Sold By Region")  # Add context in title
plot_1 <- ggplotly(plot_qty)  # Make plot interactive
plot_1

```

## Activity 2: Interactive Tables
### Steps:
- Use DT::datatable() to convert a data frame
- Add options like page length, search box, column formatting

### Example:
```{r}
# summary_tbl <- example_data %>%
#   group_by(location) %>%
#   summarize(Average_Revenue = mean(sales, na.rm = TRUE))  # Safe averaging
# datatable(summary_tbl)  # Render as interactive table

```

The code below shows the method to have interactive tables in an Rmarkdown file.

## Try it Out #1: Display a summary table of revenue per Region.
### Steps:
- Group by region and summarize average revenue
- Use datatable() to render it
- Hint: Use dplyr::summarize() to get the mean.

```{r}
summary_tbl <- sales_data %>%
  group_by(region) %>%
  summarize(Average_revenue = mean(revenue, na.rm = TRUE))

datatable(summary_tbl)

```


## Activity 3: Interactive Filters and Controls
### Steps:
- Use params in YAML or Shiny inputs like selectInput()
- Filter data based on input values

### Example
```{r}
# category_data <- example_data %>%
#  filter(category == params$category)  # Use parameter for filtering

# plot_a <- ggplot(category_data, aes(x = sales_date, y = sales)) +
#   geom_line() +
#   labs(title = paste("Sales Trend for", params$category))
# print(plot_a)
```

## Try it Out #1: Let users choose a category to explore.

## Steps:
- Add a parameter called category in YAML
- Filter data where Category == params$category
- Plot or summarize the filtered result
- Hint: Use unique(sales_data$Category) to see your options

```{r}
filtered_data <- sales_data %>% 
  filter(category == params$category)

plot_2 <- ggplot(filtered_data, aes(x = date, y = units_sold)) +
  geom_line() +
  geom_smooth(method = "lm") +
  labs(title = paste("Sales Trend for", params$category))
print(plot_2)

```

# Activity 4: Validate and Optimize Interactions

## Steps:
- Check input validity (e.g., validate that the category exists)
- Use conditional logic to avoid blank plots
- Test layout, speed, and usability

### Example:
```{r}
# if (nrow(category_data) > 0) {
#   plot_b <- ggplot(category_data, aes(x = date, y = units_sold)) +
#     geom_col(fill = "skyblue") +
#     labs(title = paste("Units Sold for", params$category))
#   print(plot_b)
# } else {
#   print("No records available for the selected category.")
# }
```

## Try it Out #1: Add fallback behavior when filters return no results.
## Steps:
- Wrap your chart or table in an if statement
- Print a message if the result is empty

```{r}
if(nrow(filtered_data) > 0) {
  plot_3 <- ggplot(filtered_data, aes(x = date, y = units_sold)) +
  geom_line() +
  geom_smooth(method = "lm") +
  labs(title = paste("Sales Trend for", params$category))
  print(plot_3)
} else {
  print("No records available for the selected category.")
}

```

## Optional: Knit your report to HTML and view it. To do this you can either:
- Run `rmarkdown::render("retail.Rmd")` in your R VSCode Terminal
- Run the `Knit RMD` button in VSCode

To view your final html document, right-click on `retail.html` and select "Open with Live Server".
This will open up your html file in your lab's internet browser. 